A performance assessment of evolutionary algorithms in volunteer computing environments

### 2016-11-13 EvoPar2017/paper.Rnw
Adds new graph refs #131

<<setup, cache=FALSE,echo=FALSE>>=
library("ggplot2")
experiment.data <- data.frame(Experiment=character(),
                              milliseconds=character(),                           
                              IPs=character(), 
                              actualIPs=character(),
                              PUTs=character(),
                              actualPUTs=character(),
                              PUTratio = character(),
                              PUTsIP =character())

@@

@@

NodIO needs a fitness function to work with; in this case the classical Trap 
function \cite{Ackley1987} has been used. % I don't think this use of the ; is correct
% Changed to .
The evolutionary algorithm that evaluates that function runs in the
browser, but the server needs to know when the solution has been
found, so a function that checks termination condition is also included in the server, as a configuration
option. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes.
For every new individual sent by the clients to the cache, several
% We have not presented the "cache",  only the server and
% ..storing the incoming chromosomes in a data structure
% Im changing over there to cache
% That's been changed - JJ
pieces of metadata were stored: a time stamp, the client IP, the chromosome and
fitness value, the cache size in that particular moment, and also if that
individual actually updated the cache or not.

@@

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, specially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph that shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

>>>>

<<setup, cache=FALSE,echo=FALSE>>=
library("ggplot2")
library("RColorBrewer")

experiment.data <- data.frame(Experiment=character(),
                              milliseconds=character(),                           
                              IPs=character(), 
                              actualIPs=character(),
                              PUTs=character(),
                              actualPUTs=character(),
                              PUTratio = character(),
                              PUTsIP =character())

@@

# New data for this paper

IPs <- read.csv("../data/nodio-2016-3-17-0_per_minute.csv")
IPs$minute <-  as.POSIXct(IPs$time, format="%Y-%m-%dT%H:%M")

minutes.data <- data.frame(Experiment="17-0",                      
                           minute=IPs$minute,
                           IPs=IPs$IPs,
                           PUTs=IPs$PUTs,
                           updates=IPs$updates,
                           ratio=IPs$updates/IPs$PUTs)

@@

NodIO needs a fitness function to work with; in this case the classical Trap 
function \cite{Ackley1987} has been used. % I don't think this use of the ; is correct
% Changed to .
The evolutionary algorithm that evaluates that function runs in the
browser, but the server needs to know when the solution has been
found, so a function that checks termination condition is also included in the server, as a configuration
option. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes.
For every new individual sent by the clients to the cache, several
pieces of metadata were stored: a time stamp, the client IP, the chromosome and
fitness value, the cache size in that particular moment, and also if that
individual actually updated the cache or not.

@@

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE,fig.height=4>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, specially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph that shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

\section{Finding the keys to a good performance}
%
%
\begin{figure}[ht!b]
\centering
<<minutedata, cache=FALSE,echo=FALSE,fig.height=6>>=
ggplot(data=minutes.data, aes(x=minute,y=PUTs,color=ratio))+ scale_x_datetime()+geom_point()+stat_smooth()
@ %def graph on PUTs per minute
%
\caption{Plotting the contributions per minute in a new and massive
  experiment, as well as the real number as color. Lighter color
  indicates that all PUTs contributed new individuals to the
  experiment, darker color few to none. \label{fig:minute}}
\end{figure}

### 2016-11-13 EvoPar2017/paper.tex
Cleaning up a bit
Adds new graph refs #131

%% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
%%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}
\author{First Author \and Second Author}

%% \institute{Dept. of Computer Architecture and Technology, University
%% of Granada, Spain \and
%% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}
\institute{First Institute in Somecountry \and Second University in Someothercountry}

@@

@@

%This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
%Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
%2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
%gathering users for the new experiments.
Acknowledgements\\
taking\\
this\\
space

>>>>

% \author{Juan-J.~Merelo \and Paloma de las Cuevas \and Pablo
%   Garc\'ia-S\'anchez \inst{1} \and Mario Garc\'ia-Valdez \inst{2}}

% \institute{Dept. of Computer Architecture and Technology, University
% of Granada, Spain \and
% Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}

\author{A.U. Thor \and A. N. Otherauthor \and Just A. Nother \inst{1}
  \and Jet A. Nother \inst{2}}

\institute{Invisible University, Frantzia \and
Institute of Higher Studies, Zootropa}

@@

\section{Finding the keys to a good performance}
%
%
\begin{figure}[ht!b]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/minutedata-1}

\end{knitrout}
%
\caption{Plotting the contributions per minute in a new and massive
  experiment, as well as the real number as color. Lighter color
  indicates that all PUTs contributed new individuals to the
  experiment, darker color few to none. \label{fig:minute}}
\end{figure}

@@

% This work has been supported in part by TIN2014-56494-C4-3-P (Spanish
% Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio
% 2015 UGR). We are also grateful to {\tt @otisdriftwood} for his help
% gathering users for the new experiments.

Several projects\\
Taking\\
This\\
Much\\
Space\\
